---
format:
  typst:
    papersize: us-letter
    margin:
      x: 1.6cm
      y: 1.4cm
    mainfont: "Helvetica Neue"
params:
  district: "District A"
  min_n: 10
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
library(patchwork)

# ======================================================================
#  BRAND CONFIG  —  edit this block to rebrand the whole one-pager.
# ----------------------------------------------------------------------
# 1) Colors: replace the placeholder hex codes with your brand colors.
#    `primary` is the header-band fill; the band text is white, so keep
#    `primary` a DARK color.
brand <- list(
  primary = "#1F4E79",   # header band + section headings (keep dark)
  ink     = "#1A1A1A",   # body text / big numbers fallback
  muted   = "#6B7280",   # captions, footnotes, source line
  track   = "#E6E8EC",   # the "rest" (un-filled) part of each donut
  rule    = "#D7DBE0",   # thin divider lines / box borders
  panel   = "#F4F6F8"    # summary box fill
)

# 2) One color per factor (the filled part of each donut).
construct_cols <- c("Interest & self-efficacy" = "#1F4E79",
                    "Belonging"                = "#2E9E8F",
                    "Classroom culture"        = "#E08A1E")

# 3) Logos: drop files in ./logos/ (PNG or SVG). Missing files are
#    skipped automatically, so the page still renders without them.
logo_top_left  <- "logos/logo_top_left.png"
logo_top_right <- "logos/logo_top_right.png"
logo_footer    <- "logos/logo_footer.png"

# 4) Summary text (the team writes the real summary here).
summary_text <- "SUMMARY GOES HERE — to be written by the team."

# Helper: emit a Typst image() call only if the file exists, else empty.
emit_logo <- function(path, height = "1.0cm") {
  if (!is.null(path) && file.exists(path)) {
    sprintf('image("%s", height: %s)', path, height)
  } else {
    "[]"
  }
}
# ======================================================================

# ---- Item dictionary: raw item -> construct -> short label -> reverse flag ----
# Edit the labels/constructs here; everything downstream is driven off this table.
# Constructs reflect the 3-factor CFA solution (see cfa_3factor.R).
items <- tribble(
  ~item,           ~construct,                   ~label,                        ~reverse,
  "enjoy",         "Interest & self-efficacy",   "Enjoy math this year",         FALSE,
  "interesting",   "Interest & self-efficacy",   "Find it interesting",          FALSE,
  "important",     "Interest & self-efficacy",   "Important to everyday life",   FALSE,
  "math_person",   "Interest & self-efficacy",   "See myself as a math person",  FALSE,
  "can_solve",     "Interest & self-efficacy",   "Can solve math problems",      FALSE,
  "belonging",     "Belonging",                  "Sense of belonging",           FALSE,
  "valued",        "Belonging",                  "Feel like a valued member",    FALSE,
  "comfortable",   "Belonging",                  "Comfortable sharing ideas",    FALSE,
  "many_ways",     "Classroom culture",          "We value many ways",           FALSE,
  "mistakes",      "Classroom culture",          "Mistakes help learning",       FALSE,
  "explain",       "Classroom culture",          "Asked to explain thinking",    FALSE,
  "classmates",    "Classroom culture",          "Classmates have good ideas",   FALSE,
  "teacher_talks", "Classroom culture",          "Teacher does most talking",    TRUE
)

construct_levels <- c("Interest & self-efficacy", "Belonging", "Classroom culture")

# ---- Load the cleaned, tidy responses (produced by prep_data.R) ----
# Long format, columns: district, school, student_id, item, response (1-4).
responses <- readr::read_csv("responses.csv", show_col_types = FALSE)

# ---- Filter to the district passed in by the render loop ----
district_data <- responses |> filter(district == params$district)
n_students    <- n_distinct(district_data$student_id)
n_schools     <- n_distinct(district_data$school)

# Fail clearly if the district filter matched nothing, instead of erroring
# deep in the plotting code.
if (nrow(district_data) == 0) {
  stop("No responses for district = '", params$district, "'. ",
       "Available districts: ",
       paste(sort(unique(responses$district)), collapse = ", "))
}

# ---- % who agree or strongly agree per item, with small-cell suppression ----
# 4-point scale (1=strongly disagree ... 4=strongly agree): "agree or
# strongly agree" = the top two boxes = response >= 3.
agree <- district_data |>
  group_by(item) |>
  summarise(n = sum(!is.na(response)),
            pct_agree = mean(response >= 3, na.rm = TRUE) * 100,
            .groups = "drop") |>
  left_join(items, by = "item") |>
  mutate(
    suppressed = n < params$min_n,
    # direction-corrected: flip reverse-coded items so that, for every
    # item, a higher % means "more student-centered / more positive".
    dir_pct    = ifelse(reverse, 100 - pct_agree, pct_agree),
    construct  = factor(construct, levels = construct_levels)
  )

# ---- One aggregate % per factor ----
construct_pct <- agree |>
  filter(!suppressed) |>
  group_by(construct) |>
  summarise(pct = round(mean(dir_pct)), .groups = "drop") |>
  complete(construct = factor(construct_levels, levels = construct_levels))
pct_lookup <- setNames(construct_pct$pct, as.character(construct_pct$construct))
```

```{r}
#| label: plot-functions
# Clean, modern donut: thin ring, big number in the factor color, the
# factor name beneath. The colored arc is the aggregate % agree.
plot_pie <- function(construct_name, pct, fill_col) {
  hsize <- 4  # larger -> bigger hole / thinner ring
  base <- theme_void(base_size = 11) +
    theme(plot.title  = element_text(face = "bold", size = 10.5,
                                     colour = brand$ink, hjust = 0.5,
                                     margin = margin(t = 6)),
          plot.margin = margin(6, 10, 6, 10))

  if (is.na(pct)) {
    return(
      ggplot() +
        annotate("text", x = 0, y = 0, label = "n too small",
                 size = 3.5, colour = brand$muted) +
        labs(title = construct_name) + base
    )
  }

  d <- tibble(part  = factor(c("agree", "rest"), levels = c("agree", "rest")),
              value = c(pct, 100 - pct))
  ggplot(d, aes(x = hsize, y = value, fill = part)) +
    geom_col(width = 1, colour = "white", linewidth = 0.6) +
    coord_polar(theta = "y") +
    xlim(c(0.2, hsize + 0.5)) +
    scale_fill_manual(values = c(agree = fill_col, rest = brand$track),
                      guide = "none") +
    annotate("text", x = 0.2, y = 0, label = paste0(pct, "%"),
             size = 9, fontface = "bold", colour = fill_col) +
    labs(title = construct_name) + base
}
```

```{r}
#| label: header
#| output: asis
cat(paste0(
'```{=typst}\n',
'#grid(\n',
'  columns: (1fr, 1fr),\n',
'  align: (left + horizon, right + horizon),\n',
'  ', emit_logo(logo_top_left), ',\n',
'  ', emit_logo(logo_top_right), ',\n',
')\n',
'#v(10pt)\n',
'#block(\n',
'  width: 100%,\n',
'  fill: rgb("', brand$primary, '"),\n',
'  inset: (x: 16pt, y: 13pt),\n',
'  radius: 6pt,\n',
')[\n',
'  #text(fill: white, weight: "bold", size: 18pt)[', params$district, ']\n',
'  #v(3pt)\n',
'  #text(fill: white, size: 10pt)[Student math perceptions  ·  Spring 2026  ·  ',
       n_students, ' students across ', n_schools, ' ',
       if (n_schools == 1) "school" else "schools", ']\n',
']\n',
'#v(14pt)\n',
'```\n'
))
```

```{r}
#| label: fig-onepager
#| fig-width: 7.6
#| fig-height: 3.1
pies <- imap(construct_cols, ~plot_pie(.y, pct_lookup[[.y]], .x))

(pies[["Interest & self-efficacy"]] | pies[["Belonging"]] | pies[["Classroom culture"]]) +
  plot_annotation(
    caption = paste0(
      "Each circle = average % who agree or strongly agree across the factor's items.  ",
      "Reverse-coded items are flipped so higher = more student-centered.  ",
      "Items below the response minimum are excluded."),
    theme = theme(plot.caption = element_text(size = 7.5, colour = brand$muted, hjust = 0))
  )
```

```{r}
#| label: summary-box
#| output: asis
cat(paste0(
'```{=typst}\n',
'#v(12pt)\n',
'#block(\n',
'  width: 100%,\n',
'  fill: rgb("', brand$panel, '"),\n',
'  stroke: 0.5pt + rgb("', brand$rule, '"),\n',
'  inset: 13pt,\n',
'  radius: 6pt,\n',
')[\n',
'  #text(weight: "bold", size: 11pt, fill: rgb("', brand$primary, '"))[Summary]\n',
'  #v(5pt)\n',
'  #text(size: 10pt, fill: rgb("', brand$ink, '"))[', summary_text, ']\n',
']\n',
'```\n'
))
```

```{r}
#| label: footer
#| output: asis
cat(paste0(
'```{=typst}\n',
'#v(14pt)\n',
'#line(length: 100%, stroke: 0.5pt + rgb("', brand$rule, '"))\n',
'#v(6pt)\n',
'#grid(\n',
'  columns: (1fr, auto),\n',
'  align: (left + horizon, right + horizon),\n',
'  text(size: 8pt, fill: rgb("', brand$muted,
       '"))[Source: Spring 2026 student math perceptions survey (grades 4–5).  Prepared ',
       format(Sys.Date(), "%B %Y"), '.],\n',
'  ', emit_logo(logo_footer, height = "0.9cm"), ',\n',
')\n',
'```\n'
))
```
