---
format:
  typst:
    papersize: us-letter
    margin:
      x: 1.3cm
      y: 1.1cm
    mainfont: "Helvetica Neue"
params:
  district: "District A"
  min_n: 10
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
library(patchwork)

# ======================================================================
#  BRAND CONFIG  —  edit this block to rebrand the whole one-pager.
# ----------------------------------------------------------------------
# 1) Colors:
brand <- list(
  primary = "#08436A",   # header band + section headings (keep dark)
  ink     = "#1A1A1A",   # body text / big numbers fallback
  muted   = "#6B7280",   # captions, footnotes, source line
  track   = "#E6E8EC",   # the "rest" (un-filled) part of each donut
  rule    = "#D7DBE0",   # thin divider lines / box borders
  panel   = "#F2C75C"    # summary box fill
)

# 2) One color per factor
construct_cols <- c("Attitudes" = "#1F4E79",
                    "Belonging"                = "#2E9E8F",
                    "Classroom Culture"        = "#E08A1E")

# 3) Logos.
logo_top_left  <- "logos/maisa_logo_int_simple.png"
logo_top_right <- "logos/early_math_logo.png"
logo_footer    <- "logos/MenloEDU-logo-Day-CMYK.png"
# Logo sizes (bump these to resize). Any Typst length works, e.g. "1.4cm".
logo_top_height    <- "1.0cm"   # the two top logos
logo_footer_height <- "1.5cm"   # the footer logo

# 4) Summary text (summary text can go here and then will render below).
summary_text <- "The Michigan Association of Intermediate School Administrators (MAISA) has been working with districts and schools to support early mathematics teaching and learning across the state of Michigan. The work supports new ways of knowing, doing, and becoming for children, teachers, and leaders. Specifically for children, the work supports a vision of mathematics learning where kids are encouraged to share and discuss ideas, experiment with multiple solutions, collaborate with classmates to figure things out, and experience joy through mathematics. As part of the external evaluation of this work, this survey measured Michigan kids' experiences of mathematics learning at school."

# Helper: 
emit_logo <- function(path, height = "1.0cm") {
  if (!is.null(path) && file.exists(path)) {
    sprintf('image("%s", height: %s)', path, height)
  } else {
    "[]"
  }
}
# ======================================================================

# ---- Item dictionary: raw item -> construct -> short label -> reverse flag ----
# Constructs reflect the 3-factor CFA solution (see cfa_3factor.R).
items <- tribble(
  ~item,           ~construct,                   ~label,                        ~reverse,
  "enjoy",         "Attitudes",   "Enjoy math this year",         FALSE,
  "interesting",   "Attitudes",   "Find it interesting",          FALSE,
  "important",     "Attitudes",   "Important to everyday life",   FALSE,
  "math_person",   "Attitudes",   "See myself as a math person",  FALSE,
  "can_solve",     "Attitudes",   "Can solve math problems",      FALSE,
  "belonging",     "Belonging",                  "Sense of belonging",           FALSE,
  "valued",        "Belonging",                  "Feel like a valued member",    FALSE,
  "comfortable",   "Belonging",                  "Comfortable sharing ideas",    FALSE,
  "many_ways",     "Classroom Culture",          "We value many ways",           FALSE,
  "mistakes",      "Classroom Culture",          "Mistakes help learning",       FALSE,
  "explain",       "Classroom Culture",          "Asked to explain thinking",    FALSE,
  "classmates",    "Classroom Culture",          "Classmates have good ideas",   FALSE,
  "teacher_talks", "Classroom Culture",          "Teacher does most talking",    TRUE
)

construct_levels <- c("Attitudes", "Belonging", "Classroom Culture")

# # Long format, columns: district, school, student_id, item, response (1-4).
responses <- readr::read_csv("responses.csv", show_col_types = FALSE)

# ---- Filter to the district passed in by the render loop ----
district_data <- responses |> filter(district == params$district)
n_students    <- n_distinct(district_data$student_id)
n_schools     <- n_distinct(district_data$school)

# Fail clearly if the district filter matched nothing, instead of errors
if (nrow(district_data) == 0) {
  stop("No responses for district = '", params$district, "'. ",
       "Available districts: ",
       paste(sort(unique(responses$district)), collapse = ", "))
}

# ---- Pooled % agree/strongly agree per factor, with small-cell suppression ----

item_n <- district_data |>
  group_by(item) |>
  summarise(n = sum(!is.na(response)), .groups = "drop") |>
  mutate(suppressed = n < params$min_n)

# Response-level positives, direction-corrected, then pooled per factor.
construct_pct <- district_data |>
  filter(!is.na(response)) |>
  left_join(select(items, item, construct, reverse), by = "item") |>
  left_join(select(item_n, item, suppressed), by = "item") |>
  filter(!suppressed) |>
  mutate(positive = ifelse(reverse, response >= 3, response <= 2),
         construct = factor(construct, levels = construct_levels)) |>
  group_by(construct) |>
  summarise(pct = round(sum(positive) / n() * 100), .groups = "drop") |>
  complete(construct = factor(construct_levels, levels = construct_levels))
pct_lookup <- setNames(construct_pct$pct, as.character(construct_pct$construct))
```

```{r}
#| label: plot-functions

plot_pie <- function(construct_name, pct, fill_col) {
  hsize <- 4  # larger -> bigger hole / thinner ring
  base <- theme_void(base_size = 11) +
    theme(plot.title  = element_text(face = "bold", size = 12,
                                     colour = brand$ink, hjust = 0.5,
                                     margin = margin(t = 3)),
          plot.margin = margin(2, 8, 2, 8))

  if (is.na(pct)) {
    return(
      ggplot() +
        annotate("text", x = 0, y = 0, label = "n too small",
                 size = 3.5, colour = brand$muted) +
        labs(title = construct_name) + base
    )
  }

  d <- tibble(part  = factor(c("agree", "rest"), levels = c("agree", "rest")),
              value = c(pct, 100 - pct))
  ggplot(d, aes(x = hsize, y = value, fill = part)) +
    geom_col(width = 1, colour = "white", linewidth = 0.6) +
    coord_polar(theta = "y") +
    xlim(c(0.2, hsize + 0.5)) +
    scale_fill_manual(values = c(agree = fill_col, rest = brand$track),
                      guide = "none") +
    annotate("text", x = 0.2, y = 0, label = paste0(pct, "%"),
             size = 9, fontface = "bold", colour = fill_col) +
    labs(title = construct_name) + base
}
```

```{r}
#| label: header
#| output: asis
cat(paste0(
'```{=typst}\n',
'#grid(\n',
'  columns: (1fr, 1fr),\n',
'  align: (left + horizon, right + horizon),\n',
'  ', emit_logo(logo_top_left, height = logo_top_height), ',\n',
'  ', emit_logo(logo_top_right, height = logo_top_height), ',\n',
')\n',
'#v(10pt)\n',
'#block(\n',
'  width: 100%,\n',
'  fill: rgb("', brand$primary, '"),\n',
'  inset: (x: 16pt, y: 13pt),\n',
'  radius: 6pt,\n',
')[\n',
'  #text(fill: white, weight: "bold", size: 20pt)[', params$district, ']\n',
'  #v(3pt)\n',
'  #text(fill: white, size: 15pt)[Student Math Survey  ·  Spring 2026  ·  ',
       n_students, ' students across ', n_schools, ' ',
       if (n_schools == 1) "school" else "schools", ']\n',
']\n',
'#v(4pt)\n',
'```\n'
))
```

```{r}
#| label: onepager
#| fig-width: 7.6
#| fig-height: 2.6
pies <- imap(construct_cols, ~plot_pie(.y, pct_lookup[[.y]], .x))

(pies[["Attitudes"]] | pies[["Belonging"]] | pies[["Classroom Culture"]]) +
  plot_annotation(
    caption = paste0(
      "Each circle represents the percent of agree and strongly agree responses across the survey items in each domain.  "),
    theme = theme(plot.caption = element_text(size = 8.5, colour = brand$muted, hjust = 0))
  )
```

```{r}
#| label: descriptors
#| output: asis
# Short descriptor under each pie, color-coded to factor. Edit the
# text here.
descriptions <- c(
  "Attitudes" = "Children's attitudes about mathematics can shape motivation, engagement and persistence and influence learning and achievement. Here, attitudes reflect how much children enjoy and value math, find it relevant beyond the classroom, and see themselves as capable math learners.",
  "Belonging" = "A sense of belonging is closely tied to children's willingness to engage and take intellectual risks in math. Here, belonging reflects children feeling like a valued class member and thier comfort with sharing ideas.",
  "Classroom Culture" = "Classroom culture sets the foundation for how children view and engage with mathematics as a discipline. Here, classroom culture includes things like valuing children's mathematical ideas, elevating multiple solution paths, and learning from mistakes."
)

desc_cell <- function(name) {
  col <- construct_cols[[name]]
  paste0(
    '[\n',
    '  #text(weight: "bold", size: 14pt, fill: rgb("', col, '"))[', name, ']\n',
    '  #v(3pt)\n',
    '  #line(length: 100%, stroke: 1.5pt + rgb("', col, '"))\n',
    '  #v(5pt)\n',
    '  #text(size: 10pt, fill: rgb("', brand$ink, '"))[', descriptions[[name]], ']\n',
    ']'
  )
}

cat(paste0(
'```{=typst}\n',
'#v(8pt)\n',
'#grid(\n',
'  columns: (1fr, 1fr, 1fr),\n',
'  column-gutter: 16pt,\n',
'  ', desc_cell("Attitudes"), ',\n',
'  ', desc_cell("Belonging"), ',\n',
'  ', desc_cell("Classroom Culture"), ',\n',
')\n',
'```\n'
))
```

```{r}
#| label: summary-box
#| output: asis
cat(paste0(
'```{=typst}\n',
'#v(8pt)\n',
'#block(\n',
'  width: 100%,\n',
'  fill: rgb("', brand$panel, '"),\n',
'  stroke: 0.5pt + rgb("', brand$rule, '"),\n',
'  inset: 13pt,\n',
'  radius: 6pt,\n',
')[\n',
'  #text(weight: "bold", size: 13pt, fill: rgb("', brand$primary, '"))[Summary]\n',
'  #v(5pt)\n',
'  #text(size: 11.5pt, fill: rgb("', brand$ink, '"))[', summary_text, ']\n',
']\n',
'```\n'
))
```

```{r}
#| label: footer
#| output: asis
cat(paste0(
'```{=typst}\n',
'#v(10pt)\n',
'#line(length: 100%, stroke: 0.5pt + rgb("', brand$rule, '"))\n',
'#v(6pt)\n',
'#grid(\n',
'  columns: (1fr, auto),\n',
'  align: (left + horizon, right + horizon),\n',
'  text(size: 12pt, fill: rgb("', brand$muted,
       '"))[Source: Spring 2026 Student Math Survey (Grades 4 & 5).  Prepared ',
       format(Sys.Date(), "%B %Y"), '.],\n',
'  ', emit_logo(logo_footer, height = logo_footer_height), ',\n',
')\n',
'```\n'
))
```
